The \project{unRAVE} catalog

### 2016-09-05 article/unrave.tex
typo in Wojno, lay out groundwork for discussion

The target selection for \rave\ is based on the $I$-band apparent magnitude,
$9 < I < 12$, with a weak $J - K_s > 0.5$ cut near the disk and bulge \citep{Wonjo_2016}.  
The $I$ band was used for the target selection because it has a good overlap with the
wavelength range that \rave\ operates in:  8410--8795~\AA.  This spectral region 
is very well-studied.  It is one of the key setups used for the ground-based 
high-resolution \ges\ survey \citep{Gilmore_2012,Randich_2013}, and the 
\rave\ resolution and wavelength coverage is comparable to the Radial Velocity 
Spectrometer on board the \gaia\ space telescope \citep{Kordopatis_2011,Recio-Blanco_2016}.  
The region includes the \ion{Ca}{2} near infrared triplet lines --- strong 
transitions that are dominated by pressure broadening --- which are visible even 
in metal-poor stars or spectra with very low signal-to-noise (S/N) ratios.
Atomic transitions of 
light-, $\alpha$-, and Fe-peak elements are also present, allowing for detailed 
chemical abundance studies.

@@

There are some discrepancies between our results and these literature studies,
and these differences arise from the manner in which we join results from the
main-sequence and giant model.  Specifically, for metal-poor stars near
the turn-off in these literature comparisons, in our weighted results we find 
these stars to have similar metallicities, but to be slightly more evolved: 
on the sub-giant branch or the base of the giant branch.  This effect is best
illustrated in Figure~\ref{fig:gold-standard-hrd}, where we show a 
Hertzsprung-Russell diagram for all stars in common between this work and that
of \citet{Bensby_2014,Reddy_2003,Reddy_2006,Valenti_Fischer_2005}.  While the
overall shape of the turn-off in our work is consistent with the literature
studies, the linked data points show that metal-poor stars in the literature
appear as more evolved stars in our combined results.  Instead, if we only
take the results from our main-sequence model --- ignoring the influence of the
giant model --- and compare those values to these literature studies, no such 
systematic discrepancies are present in the Hertzsprung-Russell diagram: we 
find those metal-poor turn-off stars to also be turn-off stars, with comparable
metallicities.  Our investigation indicates that this issue affects dwarf stars
near the turn-off that have $[{\rm Fe/H}] \lesssim -1$.  While the agreement
would improve if we adopted the main-sequence results for these specific stars,
the underlying issue of metal-poor dwarf stars would still be present.  Thus,
we use this example to illustrate caution for studies making inferences based
on main-sequence dwarf stars.  The results from both the main-sequence star model, 
and the giant star model, are available at \stub{should we put them online somewhere
for open-science, or just make them available upon request so that we don't
muddy the waters and confuse people as to which catalog to use?}

@@

\stub{Chemistry + Kinematics = Stockholm}

\stub{What do the two provide us with?}

\stub{What is one interesting science result should we show based on these abundances and Hipparcos results, which was not known previously?}

@@

\bibitem[Wonjo et al.(2016)]{Wonjo_2016} Wonjo, J., et al.\ 2016, submitted

>>>>

The target selection for \rave\ is based on the $I$-band apparent magnitude,
$9 < I < 12$, with a weak $J - K_s > 0.5$ cut near the disk and bulge \citep{Wojno_2016}.  
The $I$ band was used for the target selection because it has a good overlap with the
wavelength range that \rave\ operates in:  8410--8795~\AA.  This spectral region 
is very well-studied.  It is one of the key setups used for the ground-based 
high-resolution \ges\ survey \citep{Gilmore_2012,Randich_2013}, and the 
\rave\ resolution and wavelength coverage is comparable to the Radial Velocity 
Spectrometer on board the \gaia\ space telescope \citep{Kordopatis_2011,Recio-Blanco_2016}.  
The region includes the \ion{Ca}{2} near infrared triplet lines --- strong 
transitions that are dominated by pressure broadening --- which are visible even 
in metal-poor stars or spectra with very low signal-to-noise (S/N) ratios.
Atomic transitions of 
light-, $\alpha$-, and Fe-peak elements are also present, allowing for detailed 
chemical abundance studies.

@@

There are some discrepancies between our results and these literature studies,
and these differences arise from the manner in which we join results from the
main-sequence and giant model.  Specifically, for metal-poor stars near
the turn-off in these literature comparisons, in our weighted results we find 
these stars to have similar metallicities, but to be slightly more evolved: 
on the sub-giant branch or the base of the giant branch.  This effect is best
illustrated in Figure~\ref{fig:gold-standard-hrd}, where we show a 
Hertzsprung-Russell diagram for all stars in common between this work and that
of \citet{Bensby_2014,Reddy_2003,Reddy_2006,Valenti_Fischer_2005}.  While the
overall shape of the turn-off in our work is consistent with the literature
studies (right panel), the linked data points in the first three panels show that metal-poor stars in the literature
appear as more evolved stars in our combined results.  Instead, if we only
take the results from our main-sequence model --- ignoring the influence of the
giant model --- and compare those values to these literature studies, no such 
systematic discrepancies are present in the Hertzsprung-Russell diagram: we 
find those metal-poor turn-off stars to also be turn-off stars, with comparable
metallicities.  Our investigation indicates that this issue affects dwarf stars
near the turn-off that have $[{\rm Fe/H}] \lesssim -1$.  While the agreement
would improve if we adopted the main-sequence results for these specific stars,
the underlying issue of metal-poor dwarf stars would still be present.  Thus,
we use this example to illustrate caution for studies making inferences based
on main-sequence dwarf stars.  The results from both the main-sequence star model, 
and the giant star model, are available at \stub{should we put them online somewhere
for open-science, or just make them available upon request so that we don't
muddy the waters and confuse people as to which catalog to use?}

We note that this issue does not appear to affect our derived metallicities for these
stars.  For this reason, if stars affected by this issue can be confirmed to be
dwarfs by external information --- rather than being sub-giants as our analysis
suggests for some --- then the metallicity we report can still be considered
useful.  For the more metal-rich dwarf stars, the sample we present here
will undoubtedly be informative for upcoming studies based on exoplanet host
star properties.  At present, however, there are just $\approx$30 stars in 
\rave\ that overlap with the compilations of exoplanet host star properties 
listed at \texttt{exoplanets.org} and \texttt{exoplanets.eu}.

@@

% What information have we extracted: almost maximal use of giant branch,
% lesss so for dwarfs.

% Caveats in the results: metal-poor dwarfs/sub-giants
% HETEROGENEOUS TRAINING SET
% UNSURE IF TWO ABUNDANCE SCALES ARE ON THE SAME SCALE --!
% METAL POOR ABUNDANCES
% recommended SNR and chi-sq etc
% weird projections in space. --> use morphological classifications to guide

% despite these limitations, this catalog is the largest??? of abundances.

% in the future we will do spectrophotometric distances to improve on this

@@

\bibitem[Wojno et al.(2016)]{Wojno_2016} Wojno, J., et al.\ 2016, submitted
